---
title: "Missing Data Case Study"
output: html_notebook
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(mice))
```

Consider predicting the air quality (ozone levels) in New york mid-year. We'll use the `airquality` dataset, recorded for mid 1973:

```{r}
airquality
```

## Step 0: What data are missing?

There are some missing data. Use `md.pattern` to see patterns in missingness:

```{r, fig.height=3.5}
md.pattern(airquality)
```

Fill in the following:

- There are **FILL_THIS_IN** rows of complete data.
- There are **FILL_THIS_IN** rows where only ozone is missing.
- There are **FILL_THIS_IN** rows where both ozone and Solar.R are missing.
- There are **FILL_THIS_IN** rows missing an ozone measurement.
- There are **FILL_THIS_IN** `NA`'s in the dataset.

## Step 1: Handling Missing Data

### Any Ideas?

Here is a scatterplot of `Solar.R` and `Ozone`, with missing values "pushed" to the intercepts:

```{r, fig.height=3}
airquality %>% 
	mutate(missing = if_else(is.na(Solar.R) | is.na(Ozone), TRUE, FALSE),
		   Solar.R = ifelse(is.na(Solar.R), 0, Solar.R),
		   Ozone   = ifelse(is.na(Ozone), 0, Ozone)) %>% 
	ggplot(aes(Solar.R, Ozone)) +
	geom_hline(yintercept = 0, linetype = "dashed") +
	geom_vline(xintercept = 0, linetype = "dashed") +
	geom_point(aes(colour = missing)) +
	theme_bw() +
	scale_colour_discrete(guide = FALSE)
```

Discussion: What are some ways of handling the missing data? What are the consequences of these ways?


### `mice`

First, remove the day and month columns (we won't be using them):

```{r}
airquality <- airquality %>% 
	select(-Month, -Day)
```

Make `m` random imputations using `mice::mice()`:

```{r}
m <- 10
(init <- mice(airquality, m = m, printFlag = FALSE))
```

Check out the first imputed data set using `mice::complete()`. **WARNING**: there's also a `tidyr::complete()`! Rows 5 and 6, for example, originally contained missing data.

```{r}
mice::complete(FILL_THIS_IN, FILL_THIS_IN)
```

Plot one of them:

```{r, fig.height=3}
mice::complete(init, 1) %>% 
	mutate(missing = if_else(is.na(airquality$Solar.R) | 
							 	is.na(airquality$Ozone), TRUE, FALSE)) %>% 
	ggplot(aes(Solar.R, Ozone)) +
	geom_point(aes(colour = missing)) +
	theme_bw()
```

Now, fit a linear model on each data set using the `with()` generic function (method `with.mids()`:

```{r}
(fits <- with(FILL_THIS_IN, lm(Ozone ~ Solar.R + Wind + Temp)))
```

Looks can be deceiving. This is not actually a list of length `m`! Unveil its true nature:

```{r}
unclass(fits) %>% 
	str(max.level = 1)
```

It's now easier to find the `lm` fit on the first dataset:

```{r}
fits$FILL_THIS_IN[[1]]
```

Or, we can obtain a summary of each fitted model:

```{r}
summary(fits)
```

As an aside, let's demonstrate that we can also use `mice` to fit GLM's:

```{r}
with(init, glm(
	Ozone ~ Solar.R + Wind + Temp, 
	family = Gamma(link="log")
)) %>% 
	summary()
```

## Step 3: Pool results

The last step is to pool the results together:

```{r}
pool(FILL_THIS_IN)
```

The `estimate` column you see are just the averages of all `m` models.

Column names make more sense in light of the book "Multiple Imputation for Nonresponse in Surveys" by Rubin (1987), page 76-77:

- `estimate`: the average of the regression coefficients across `m` models.
- `ubar`: the average variance (i.e., average SE^2) across `m` models.
- `b`: the sample variance of the `m` regression coefficients.
- `t`: a final estimate of the SE^2 of each regression coefficient. 
	- = `ubar + (1 + 1/m) b`
- `df`: the degrees of freedom associated with the final regression coefficient estimates.
	- An `alpha`-level CI: `estimate +/- qt(alpha/2, df) * sqrt(t)`.
- `riv`: the relative increase in variance due to randomness. 
	- = `t/ubar - 1`

