---
title: "Mixed Effects Models Case Study"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(Lahman))
```

We will use the Baseball dataset `Teams` available in the `Lahman` package in R, with teams having at least 5 years on record.

```{r}
dat <- Teams %>% 
    count(teamID) %>% 
    right_join(Teams, by = "teamID") %>% 
    filter(n >= 5) %>% 
    select(team = teamID, runs = R, hits = H, walks = BB) %>% 
    mutate(hits = scale(hits),
           walks = scale(walks))
head(dat)
```

We will investigate how hits and walks affects total number of runs (data are scaled to improve convergence of model-fitting). 

Here is a plot on the effect of walks on runs:

```{r, warning=FALSE, fig.width=10}
ggplot(dat, aes(walks, runs)) +
    facet_wrap(~team, ncol=7) +
    geom_point(alpha=0.5) +
    geom_smooth(method="lm", se=FALSE, 
                fullrange=TRUE, size=1) +
    theme_bw()
```

### Two extremes: pooling and separating

Use a linear model to determine the effect of `hits` and `walks` on `runs`, on an average `team`. Best to allow each team to have its own intercept. Note the standard error on `hits` and `walks`.

```{r}
lm(runs ~ FILL_THIS_IN, data = dat) %>% 
    tidy() %>% 
    head()
```

How would you determine the effect of `hits` and `walks` _for each individual team_? Note the standard error on `hits` on a representative team.

```{r}
lm(runs ~ FILL_THIS_IN, data = dat) %>% 
    tidy() %>% 
    filter(str_detect(term, "hits"))
```

Which model yields a smaller SE?

"Pooled" model: artificially small, since data are actually correlated.

"Separate" model: artificially large, since we're only using data within a particular group.

### An intermediate: mixed effects models

Fit a mixed effects model with random effects on both `hits` and `walks`. Note the SE's are _in between_ the "pooled" and "separate" linear models.

```{r}
fit_lme <- lmer(runs ~  FILL_THIS_IN, data = dat)
tidy(fit_lme)
```

Obtain estimates on individual teams:

```{r}
coef(fit_lme)[[1]]
```

Extract the covariance matrix of the fixed effects:

```{r}
vcov(fit_lme)
```

Extract the covariance matrix of the random effects:

```{r}
summary(fit_lme)$varcor[[1]]
```

### Questions

1. Order the models in terms of increasing bias (of regression coefficients)
2. Order the models in terms of increasing variance (of regression coefficients).
3. What is the effect of `hits` on `runs` for team ARI, using the LME?
4. What is the exact distribution of effects (slopes) of `hits` on `runs`?
5. What is the SE of the fixed effect of `hits`?
6. A new team has an average number of hits and walks. Their expected number of runs has what exact distribution?

