---
title: "Care and Value of Model Assumptions: Case study"
output: html_document
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
```

Case study: how do features of nesting female horseshoe crabs influence the number of males found nearby?

Load the data. Here are the top six rows of 173 rows:

```{r}
crab <- read_table("https://newonlinecourses.science.psu.edu/stat504/sites/onlinecourses.science.psu.edu.stat504/files/lesson07/crab/index.txt", col_names = FALSE) %>% 
  select(-1) %>% 
  setNames(c("colour","spine","width","weight","n_male")) %>% 
  mutate(colour = factor(colour),
         spine  = factor(spine))
knitr::kable(head(crab))
```

Predictors: Colour; spine condition; carapace width; weight. 

First, let's see how carapace width influences the mean number of males nearby.

```{r, fig.width=6, fig.height=3}
p <- ggplot(crab, aes(width, n_male)) + 
  geom_point(alpha=0.25) +
  labs(x = "Carapace Width", 
       y = "No. males\nnearby") +
  theme_bw() +
  theme(axis.title.y = element_text(angle=0, vjust=0.5))
plotly::ggplotly(p)
```

Data source: [H. Jane Brockmann's 1996 paper](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1439-0310.1996.tb01099.x); found online [here](https://newonlinecourses.science.psu.edu/stat504/sites/onlinecourses.science.psu.edu.stat504/files/lesson07/crab/index.txt); another regression demo with this data is found [here](https://newonlinecourses.science.psu.edu/stat504/node/169/).


## Approach 1: Estimate regression curve / model function locally

### Preliminary questions

These questions are meant to check your understanding of local regression.

What is the estimated mean number of nearby males for nesting females having a carapace width of 32.5? Use the following methods, by hand.

1\. kNN with $k=3$.

```{r}

```

2\. Using a moving window with a radius of 2.4.

```{r}

```

3\. Using a kernel smoother with Gaussian kernel with variance 1.

```{r}

```

4\. Using local polynomials with a radius of 2.4 and a flat kernel, first with degree 1, then with degree 2.

```{r}

```

### Fit a smoother by eye

Optimize the loess fit by-eye. Just modify span, to keep things simple.

```{r, fig.width=6, fig.height=4}
grid <- seq(min(crab$width), max(crab$width), length.out=100)
grid_df <- tibble(width = grid)
# FIT_MODEL_HERE
# PLOT_CURVE_HERE
```

What's the error of this model? Training error is fine.

```{r}

```

How well does this model answer our original question?

## Approach 2: Linear Regression

### Fit a linear regression model

Fit a linear regression model. What's the error?

```{r}

```

How well does this model answer our original question? Do you see a potential problem with this model? Are any assumptions of linear regression not true? Brainstorm ideas for how to deal with the problems.

## Approach 3: Link Function

Fit a GLM. What's the error?

```{r}

```

